---
title: "STA 141 HW 4"
author: "Jiewei Chen (999 494 235)"
date: "11/26/2016"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# 1. Extracting Information

```{r HW 4 Question 2 as a function, echo=T, message=F, warning=F, fig.width=5.5, fig.height = 4.5, fig.align='center'}
# This HW4.q2 function returns the result of Homework 4 Question 2
# It stores the result into a csv file called "Question 2.csv"
HW4.q2 = function () {
        library(RCurl)
        library(XML)
        
        # load website
        dplink = "http://anson.ucdavis.edu/~mueller/cveng13.html"
        
        # Parses an XML or HTML file or string containing XML/HTML content, 
        # and generates an R structure representing the XML/HTML tree
        Muller = htmlTreeParse(dplink) 
        
        # names(Muller)
        # names(Muller$children) # html
        # names(Muller$children$html) # head body
        # Since after examing the webpage, there is no information in the "head".
        # So I only extract the information from "body".
        
        Muller.body = Muller$children$html[[2]]
        # class(Muller.body)
        # names(Muller.body) # div
        # length(Muller.body) # 1
        # Muller.body[[1]]
        
        # Information needed
        # ==================
        # 1. Year of Publication        ----
        # 2. Authors                    ----===> These are stored in $text
        # 3. Title of Publication       ---- 
        # 4. Journal title              ===> This is stored in $em
        # 5. Journal Volume             ===> This is in $strong
        # 6. URL for the publication    ===> $a
        
        result = sapply(1:length(Muller.body[[1]]), 
                        function(i) xmlChildren(Muller.body[[1]][[i]]))
        # ====================================================================
        # extract $text
        result.text = sapply(1:length(result), function(i) xmlValue(result[[i]]$text))
        # length(result.text) # 280, there are NA's
        # remove NA's
        index = which(is.na(result.text)==F)
        result.text = result.text[index]
        
        # ====================================================================
        # extract Year
        library(stringr)
        yr.ptn = "\\(\\d{4}\\)"
        result.year = str_extract(str_extract(result.text, yr.ptn), "\\d{4}")
        
        # ====================================================================
        # extract authors
        author.ptn = " [[:alnum:] ,\\.]*\\("
        pt = "\\s{2,}"
        # author
        result.author = str_split_fixed(result.text, "\\(", n = 2)[,1]
        # remove redundent spaces and \n
        result.author = str_trim(gsub(pt, " ", gsub("\n", " ", result.author))) 
        # remove word "and"
        result.author = gsub("and", "", result.author) 
        #  list containing co_author names
        co_authors = sapply(1:length(result.author), 
                            function(i) {str_split(result.author[i], "\\.\\,\\s")})
        # list of number of co-authors
        nm_co_authors = sapply(co_authors, length)
        
        # ====================================================================
        # extract title
        result.title = str_split_fixed(result.text, "\\)[,\\.]", n = 2)[,2]
        # remove redundent spaces and \n
        result.title = str_trim(gsub(pt, " ", gsub("\n", " ", result.title)))
        
        # ====================================================================
        # extract $em which containts Title of Journal
        result.em = sapply(1:length(result), function(i) xmlValue(result[[i]]$em))
        result.em = result.em[index]
        # formatting Journal name
        # remove redundent spaces and \n
        result.em = gsub(pt, " ", gsub("\n", " ", result.em)) 
        result.em = gsub("(Journal of (the)?)", "J.", result.em)
        result.em = gsub("\\&", "and", result.em)
        # dealing with special cases
        result.em = ifelse((result.em == "J. Royal Statistical Society"), 
                           "J. Royal Statistical Society B", 
                           ifelse( (result.em == "Proc. National Academy of Sciences USA"), 
                                   "Proceedings of the National Academy of Sciences USA",
                                   ifelse( (result.em ==  "Sankhya"), "Sankhya A", 
                                           ifelse( (result.em %in% 
                                            c("J. of Gerontology -- Biological Sciences", 
                                              "J. of Gerontology --Biological Sciences")), 
                                           "J. of Gerontology - Biological Sciences", result.em))
                                   )
        )

        
        # ====================================================================
        # extract $strong which contains "Volume"
        result.strong = sapply(1:length(result), function(i) xmlValue(result[[i]]$strong))
        result.strong = result.strong[index]
        # deal with special case
        result.strong = ifelse( (result.strong == "B 76"), "76", result.strong)
        
        # ====================================================================
        # extract $a / websites
        ### One problem here is that it cannot work with multiple a tags
        result = result[index]
        result.website = sapply(1:length(result), function(i) {
                ifelse(is.null(result[[i]]$a), NA, 
                               xmlAttrs(result[[i]]$a)) } )
        result.website[grepl("http(s)?://", result.website)==F] = NA
        
       
        # ====================================================================
        Result = as_data_frame(cbind(result.year, result.author, nm_co_authors, result.title, 
                                     result.em, result.strong, result.website))
        # Add Human readable Column names
        colnames(Result) = c("Year of Publication", "Authors", "# of Co-authors",
                             "Title of Publication", "Journal title", 
                             "Journal Volume", "URL for the publication")
        
        # removing no publication rows
        index2 = which(is.na(Result$'Year of Publication')==F) 
        Result = Result[index2,]
        
        setwd("~/Dropbox/Course Fall 2016/STA 141/HW 4")
        write_csv(Result, path = "Question2.csv", na = "NA")

}
```

```{r result, echo=T, tidy = TRUE, message=F, warning=F, fig.width=5.5, fig.height = 4.5, fig.align='center'}
library(tidyverse)
setwd("~/Dropbox/Course Fall 2016/STA 141/HW 4")
HW4.q2()
TEST.Data = read_csv("Question2.csv")
```



# 2. A Statistical Summary

```{r test, echo=T, tidy = TRUE, message=F, warning=F, fig.width=5.5, fig.height = 4.5, fig.align='center'}
dim(TEST.Data)
```

There are in total `r dim(TEST.Data)[1]` publications listed on the website.

## 2.1 Year of Publication

```{r 2.1, echo=T, message=F, warning=F, fig.width = 9, fig.height = 6, fig.align='center'}
library(ggplot2)
ggplot(TEST.Data, aes(x = TEST.Data$"Year of Publication", 
                      fill = as.factor(TEST.Data$"Year of Publication"))) +
        geom_bar(width = 0.5) + ggtitle("Number of Publications in Years") + xlab("Years") +
        ylab("Number of Publications") + guides(fill = guide_legend(title = "Years"))
# Cut Year of Publication into levels
# These levels are the same as the levels shown on the website
TEST.Data$Y2 = cut(TEST.Data$`Year of Publication`, 
                   breaks = c(1978,1989,1994, 1999, 2004, 2009, 2014, 2016), 
                   labels = c("1979-1989", "1990-1994", "1995-1999", "2000-2004", "2005-2009",
                              "2010-2014","2015-2016"))
ggplot(TEST.Data, aes(x = Y2, fill = Y2)) + geom_bar() + 
        ggtitle("Number of Publications in Years") + xlab("Years") + 
        ylab("Number of Publications") + guides(fill = guide_legend(title = "Year Range"))
```

From these two plot of "Number of Publications in Different Years", it can be found that Porf. Muller published the most paper in 2005. In the year ranging from 2005 to 2009, he published more paper than other year range.

## 2.2  Authors

### 2.2.1 Check if there is any missing values?

```{r 2.2-1, echo=T, tidy = TRUE, message=F, warning=F, fig.width = 8, fig.height = 5, fig.align='center'}
sum(is.na(TEST.Data$Authors)==T)
```

No missing value.

### 2.2.2 Finding co-authors

```{r 2.2-2, echo=T, tidy = TRUE, message=F, warning=F, fig.width = 8, fig.height = 5, fig.align='center'}
co_authors = sapply(1:length(TEST.Data$Authors),function(i) {str_split(TEST.Data$Authors[i], "\\.\\,\\s")})
# If we want to extract all the people has been co-authored with, here is the method
all.author = ifelse(grepl("[\\.]$", unlist(co_authors)), unlist(co_authors), paste0(unlist(co_authors), "."))
# Creating a list containing co_author names
all.author.list = sort(unique(all.author))
length(all.author.list)
head(all.author.list)
tail(all.author.list)
```

There are `r length(all.author.list)` co-authors shown on this website. But we must be caucious about this result, because people with the same last name and initial first name will be categorized as the same person here.

```{r 2.2_3, echo=T, tidy = TRUE, message=F, warning=F, fig.width = 8, fig.height = 5, fig.align='center'}
sort(table(all.author), decreasing = T)[1:6]
```

And it can be found that apart from Prof. Muller, the top five co-authors are "Wang, J.L.", "Carey, J.R.", "Liedo, P.", "Yao, F." and "Chen, K.".

```{r 2.2-3, echo=T, message=F, warning=F, fig.width = 8, fig.height = 5, fig.align='center'}
ggplot(TEST.Data, aes(x = TEST.Data$`# of Co-authors`, 
                      fill = factor(TEST.Data$`# of Co-authors`))) + 
        geom_bar() + ggtitle("Number of Co-authors") + xlab("Number of Co-authors") + 
        ylab("Frequency") + guides(fill = guide_legend(title = "# of Co-authors"))
```

This plot above shows that the number of co-authors in one paper is usually less than 10. Most of the paper are co-authored within two people.

## 2.3 Title of Publication

```{r 2.3, echo=T, tidy = TRUE, message=F, warning=F, fig.width = 8, fig.height = 5, fig.align='center'}
sum(is.na(TEST.Data$`Title of Publication`)==T)
```

No missing value.

## 2.4 Journal Title

### 2.4.1 Check if there is missing value?

```{r 2.4.1, echo=T, tidy = TRUE, message=F, warning=F, fig.width = 8, fig.height = 5, fig.align='center'}
sum(is.na(TEST.Data$`Journal title`)==T) 
```

There are 6 missing values out of 245.

```{r 2.4.1.1, echo=T, tidy = TRUE, message=F, warning=F, fig.width = 8, fig.height = 5, fig.align='center'}
which(is.na(TEST.Data$`Journal title`)==T)
```

For these publication, the title of Journal is unclear, so they are left as blank or NA in the result.


### 2.4.2 List of All Journal

```{r 2.4.2, echo=T, tidy = TRUE, message=F, warning=F, fig.width = 8, fig.height = 5, fig.align='center'}
sort(table(TEST.Data$`Journal title`), decreasing = T)[1:3]
# There are some Journal names started with J. To match these with other Journal 
#       with full names, so I reformatted the name for later statistical analysis.
#       See the HW4.Q2 function.
length(table(TEST.Data$`Journal title`))
```

It can be found in the journal title list that there are two duplicate names. "J. Royal Statistical Society B" and "J. Royal Statistical Society". "Proc. National Academy of Sciences USA" and "Proceedings of the National Academy of Sciences USA". "Sankhya A" and "Sankhya". "J. of Gerontology -- Biological Sciences", "J. of Gerontology --Biological Sciences" and "J. of Gerontology - Biological Sciences". I changed them manually.

There are in total `r length(table(TEST.Data$"Journal title"))` Journals where Prof. Muller has published paper on.

The top three journal where Prof. Muller has most publications are "J. American Statistical Association", "Annals of Statistics", and "Biometrika".

Below lists all the Journal titles.

```{r 2.4.2-2, echo=T, tidy = TRUE, message=F, warning=F, fig.width = 8, fig.height = 5, fig.align='center'}
sort(unique(TEST.Data$`Journal title`))
```

## 2.5 Journal Volume

### 2.5.1 Check if there is missing value?

```{r 2.5.1, echo=T, tidy = TRUE, message=F, warning=F, fig.width = 8, fig.height = 5, fig.align='center'}
sum(is.na(TEST.Data$`Journal Volume`)==T)
which(is.na(TEST.Data$`Journal Volume`)==T)
```

There are 29 missing values out of 245. After examing the excel output, it is found that there is one publication whose volume is "B 76". By looking at the website HTML script, it can be found that it is a mistake in the script. So I manually changed the excel output.

## 2.6 URL of Publication

### 2.6.1 Check if there is missing value?

```{r 2.6.1, echo=T, tidy = TRUE, message=F, warning=F, fig.width = 8, fig.height = 5, fig.align='center'}
sum(is.na(TEST.Data$`URL for the publication`)==T)
```

There are `r sum(is.na(TEST.Data$"URL for the publication")==T)` publications, which don't have URL link. 

### 2.6.2 Another Method

```{r 2.6.2, echo=T, tidy = TRUE, message=F, warning=F, fig.width = 8, fig.height = 5, fig.align='center'}
# There is another method to extract all the links from this website
library(xml2)
doc = read_html("http://anson.ucdavis.edu/~mueller/cveng13.html")
Link = xml_find_all(doc, "//a")
links = xml_attr(Link, "href")
links = links[grepl("http(s)?://", links)]
length(links) # 132
# Through this method, more links can be found. 
# Comparing the difference
setdiff(links, TEST.Data$`URL for the publication`)
```

It can be found that the first and last link are not corresponded to any publication. [3] is another version of the publication. The rest are supplement information or appendix for correspongidng paper, in my result, these are not included.

